---
title: "CSV to Parquet Quickstart"
format: html
---

# Goal

This quickstart shows a stable approach for converting CSV to Parquet when some
integer values may exceed Arrow int32 limits.

We use three modular pieces:

- `check_big_integer()`: checks whether integer-like values exceed int32.
- `generate_csv_arrow_schema()`: infers a safe Arrow schema, one column at a time.
- `convert_csv_to_parquet_pipe()`: simple pipe-based converter using Arrow.

# Why this approach

Large CSV files can be memory-heavy if all columns are loaded as strings at once.
The schema generator scans one column at a time and uses chunked Arrow reads.
This is generally more stable for large files.

# Setup

```{r}
source("csv_to_parquet_with_int64_fix.R")
```

# 1) Check big integers

```{r}
check_big_integer(c("1", "2", "2147483648"))
```

# 2) Generate schema

```{r}
schema_info <- generate_csv_arrow_schema("data/int_overflow_example.csv")
schema_info$overflow_columns
schema_info$schema
```

# 3) Convert with pipe-based function

```{r}
out <- convert_csv_to_parquet_pipe(
  csv_path = "data/int_overflow_example.csv",
  parquet_path = "data/int_overflow_example_modular.parquet",
  schema = schema_info$schema
)

out
```

# One-call safe conversion

```{r}
result <- convert_csv_to_parquet_int64_safe("data/int_overflow_example.csv")
result$overflow_columns
result$parquet_path
```

# Notes

- Integer-like columns over int32 are promoted to `int64`.
- Non-numeric columns remain `utf8`.
- Numeric columns with decimals become `float64`.
